---
title: "Accessing Wikimedia Pageviews Data"
author: "Oliver Keyes"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Accessing Wikimedia Pageviews Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The Wikimedia websites (such as Wikipedia) are visited by hundreds of millions of people a year, and so the open datasets of pageviews
contain useful information on the subjects that interest people around the globe.

The Wikimedia Foundation recently launched [a new API](https://wikimedia.org/api/rest_v1/?doc#!/Pageviews_data/) for this data that is officially supported, allows a distinction between
different types of users and different types of traffic. The `pageviews` package serves as a client for that API.

## Per-article data

The most granular data available is on a per-article basis, and can be accessed with `article_pageviews`. This takes a project,
in the form `language`.`project_class`, an article title (with or without spacing), start and (optionally) end dates, specified as R date objects, and (should you choose) the platform and user type to return. By default, it reformats the resulting data as a data.frame:

```{r}
library(pageviews)
str(article_pageviews(project = "de.wikipedia", article = "R_(Programmiersprache)"
  , start = as.Date('2015-11-01'), end = as.Date("2015-11-02")
  , user_type = c("user", "bot"), platform = c("desktop", "mobile-web")))
```

With the `user_type` and `platform` arguments you can include or exclude spiders and other automata, and switch between overall pageviews, pageviews to the desktop site, or pageviews to the mobile site or app. Both `user_type` and `platform` can be character strings or vectors, in which case all combinations of `user_type` and `platform` will be queried and returned.

## Per-project data

Per-project data can also be retrieved using (you've guessed it) `project_pageviews`. This looks very similar to `article_pageviews`
with one major difference: you can specify the granularity of the data, returning either daily or *hourly* data. The platform, user type and timestamp options are identical.

```{r}
str(project_pageviews())
```

## Top articles

With `top_articles` you can get data about the top 1,000 articles by pageviews on a project within a given timeframe (and on a given platform). This uses distinct year/month/day arguments, and for month and day, you can select "all" indicating that you want the top
articles overall in a given year or month respectively.

```{r}
str(top_articles())
```

The underlying data source here is likely to change, because it currently contains some automata, so be aware that there may be noise
or unexpected entries in data from this function.

## Bugs and feature suggestions
If there's something missing in the client that's supported by the API proper, please [submit an issue](https://github.com/Ironholds/pageviews/issues)! And, while the author of this package doesn't maintain the API, if you see functionality that you need that *isn't* covered by the API, you can still submit an issue on the client repo - I'll just courier them over to the Wikimedia bug-tracker.
